Bayesian Sparsification Methods for Deep Complex-valued Networks

### 2020-06-21 paper/paper.tex
6  clearer story in the CVNN section  * what CVNN are and how they work  * how to back-propagate though them  * and what were the major developments in their niche

% Although $\cplx$-valued dense layers and back-propagation have been around since early
% 1990-s, one of the earliest accounts

Development of deep $\cplx$-valued networks has been active. \citet{haensch_complex-valued_2010}
put forward $\cplx$-valued convolutional networks, \citet{guberman_complex_2016} and
\citet{popa_complex-valued_2017} developed modifications of pooling, \citet{arjovsky_unitary_2016}
and \citet{wisdom_full-capacity_2016} proposed $\cplx$-valued RNNs with unitary recurrent
transition matrices, and \citet{danihelka_associative_2016} developed $\cplx$-valued
holographic representations for LSTMs. More recently \citet{trabelsi_deep_2018} proposed
$\cplx$-valued batch-normalization and weight initialization, \citet{wolter_complex_2018}
investigated different $\cplx$-valued gating mechanisms for RNNs, \citet{yang_complex_2019}
proposed $\cplx$-valued self-attention and complex transformer architecture. It merits
noting that \citet{gaudet_deep_2018} generalized $\cplx$VNN further to deep quaternion-valued
networks, and \citet{vecchi_compressing_2020} studied sparsity inducing regularizers for
them.

% Nonlinearities used in $\cplx$VNN are of two classes: split that operate on equivalent
% planar or polar representations of $\cplx$, of full which consider the complex number
% as a whole. Split planar $
%   z \mapsto g(\Re{z}) + \iu g(\Im{z}) % split
% $, \citep{benvenuto_complex_1992,hirose_generalization_2012,haensch_complex-valued_2010},
% or polar $
%   z \mapsto g(\lvert z\rvert) e^{\iu \arg{\!(z)}}
% $, \citep{wolter_complex_2018,arjovsky_unitary_2016}. Full trigonometric, holomorphic, 
% % \citet{hirose_generalization_2012} polar (tanh)
% % \citet{savitha_2011} trig (sinh)
% % \citet{arjovsky_unitary_2016} polar (modReLU)
% % \citet{guberman_complex_2016} polar (zRelU)
% % \citet{wolter_complex_2018} gating for $\cplx$-RNN

% % complex-valued distributions \citep{pav_moments_2015,taubock_complex-valued_2012},
% % and \citep{karseras_caution:_2014}

Complex-valued networks are, in general, non-holomorphic, i.e. not $\cplx$-differentiable, which is
exacerbated by the loss being real-valued. Indeed a $\cplx$-differentiable real-valued function is
necessarily trivial by Cauchy-Riemann conditions. This issue is dealt with by employing Wirtinger,
or $\cplx\real$ calculus, which generalizes holomorphic calculus to non-holomorphic functions
of $\cplx$-argument \citep{adali_complex-valued_2011,boeddeker_computation_2019}. It satisfies
the product and chain rules, respects complex conjugation and linearity, but most importantly the
$\cplx\real$-differential of a function on $\cplx$ coincides with its classical differential as
a function on $\real^2$. Essentially, it allows straightforward retrofiting of $\cplx$-valued
networks into existing $\real$ deep learning auto-differentiation frameworks. It was considered
as a basis for $\cplx$ version of back-propagation by \citet{benvenuto_complex_1992}, (see
appendix~\ref{sub:wirtinger_calculus}).
% \todo{read on holomorphic nets}

More recently general development of $\cplx$-valued networks has been continued by
\citet{trabelsi_deep_2018}, who outline the essential building blocks for deep $\cplx$-valued
networks and describe suitable representation and operations, including convolutional and
dense layers, $\cplx$-valued activations, complex batch-normalization and weight initialization.
Their approach coupled with $\cplx\real$ calculus for backprop transforms a $\cplx$-valued
network into an intricately connected $\real$-valued computational graph, constructed
to respects $\cplx$-arithmetic.
%
In particular, natural identification of $\cplx$ and $\real^2$ implies that dense linear
layers $
  L \colon \cplx^m \to \cplx^n
$ act upon their inputs as follows:
\begin{equation}  \label{eq:cplx-lin-op}
  {(\real^n)}^{2}
    \to {(\real^m)}^{2}
    \colon (u, v)
      \mapsto \Bigl(
        P u - Q v,
        P v + Q u
      \Bigr)
    \,,
\end{equation}
where $
  P, Q \colon \real^{n} \to \real^{m}
$ are unique operators such that $L = P + j Q$. A $\cplx$-convolutional layer with kernel
$W$ can be implemented as two $\real$-convolutions with kernels $\Re{W}$ and $\Im{W}$ using
\eqref{eq:cplx-lin-op}. Activations include trigonometric functions and representation-dependent
non-linearities \citep{hirose_complex-valued_2009}: polar (phase-amplitude) $
  z \mapsto \sigma(\lvert z \rvert) e^{j \arg z}
$ or planar (real-imaginary) $
  z \mapsto \sigma(\Re z) + j \sigma(\Im z)
$ for some $\real$-valued non-linearity $\sigma$, possibly with learnable parameters
\citep{trabelsi_deep_2018,wolter_complex_2018}.

>>>>

% implementation
$\cplx$-valued neural networks are networks that rely on the arithmetic in the complex
domain. To achieve this implementations of $\cplx$VNN use the geometric (planar) representation
of a complex number as paired \emph{real} and \emph{imaginary} values, $\cplx \simeq \real^2$,
ensuring that the resulting $\real$-valued computational graph respects $\cplx$-arithmetic.
For example, $
  f\colon \cplx^n \to \cplx^m
$ is identified with a real vector-valued function $
  F\colon \real^{2 m} \to \real^{2 m}
$ defined via $
  F(u, v) = (\Re f(u + \iu v), \Im f(u + \iu v))
$, $\Re$ and $\Im$ denoting the real and imaginary parts, respectively. When $f$ is
a $\cplx$-valued linear transformation, the computations are ``wired'' so that
\begin{equation}  \label{eq:cplx-lin-op}
  % F\colon
  %   (u, v)
  %     \mapsto \Bigl(
  %       P u - Q v,
  %       P v + Q u
  %     \Bigr)
  F(u, v)
    =
    \begin{pmatrix}
      P u - Q v \\
      P v + Q u
    \end{pmatrix}
    = \begin{pmatrix}
      P & - Q \\ Q & P
    \end{pmatrix} \begin{pmatrix}
      u \\ v
    \end{pmatrix}
    \,,
\end{equation}
with $
  P, Q \colon \real^{n} \to \real^{m}
$ given by $P = \Re f$ and $Q = \Im f$ restricted to $\real^{n}$.
% The downside is that, for instance, one $\cplx$-valued convolution requires up to
% four $\real$-valued convolution operations.
%
Nonlinearities in $\cplx$VNN can be hyperbolic functions or maps that operate on
$\cplx$ numbers in planar form, $
  z \mapsto \sigma(\Re{z}) + \iu \sigma(\Im{z}) % split
$, or polar form $
  % z \mapsto \sigma(\lvert z\rvert, \arg{\!(z)})
  r e^{\iu \phi} \mapsto \sigma(r, \phi)
$.
% \citep{savitha_2011,hirose_generalization_2012,arjovsky_unitary_2016,guberman_complex_2016}

This $\cplx \simeq \real^2$ identification allows straightforward retrofitting of $\cplx$VNN
into existing $\real$-valued auto-differentiation frameworks for deep learning. This
act is backed by Wirtinger ($\cplx\real$) calculus, which enables generalized treatment
of functions of complex argument, by regarding $z$ and $\conj{z}$ as independent variables
and defining derivatives with respect to them through partial derivatives with respect
to real and imaginary parts of each argument. These definitions simplify manual calculations
of $\cplx$ derivatives and satisfy the product and chain rules, respect complex conjugation
and linearity for $\cplx \to \cplx$ maps. $\cplx\real$ calculus was used to define $\cplx$
version of back-propagation, \citep{benvenuto_complex_1992,guberman_complex_2016}. However,
since auto-differentiation frameworks can algorithmically handle computational graphs
of arbitrary complexity, explicit use of Wirtinger derivatives is not required, especially
considering the fact that the direction of the steepest ascent of a $\cplx\to \real$
function is given by complex conjugate gradient $\nabla_{\conj{z}}$, which coincides
with the classical gradient of the same function viewed as $\real^2 \to \real$,
(see appendix~\ref{sub:wirtinger_calculus}).
%
% in natively $\cplx$-valued backprop it would be necessary to modify the existing sgd
% to explicitly use $\nabla_{\conj{z}} f(z)$, while no such modification is necessary
% in re-im. If some inner activation of a $\cplx$VNN is holomorphic, i.e. satisfies
% Cauchy-Riemann conditions, then that particular step of in the chain rule that is
% backprop can be simplified by assuming one piece of the tracked gradient to be exact
% zero. But other wise it is necessary to track both $\partial_{\conj{z}}$ and $\partial_{z}$.
%
% Training a $\cplx$VNN involves optimizing a real-valued objective with respect to
% complex valued arguments.
%
% The computed gradients are theoretically justified
% - CR dictates the it is necessary to track both $z$ and $\conj{z}$ derivatives
% - the direction of largest increase in f: C-R is given by the gradient with
% respect to the conjugate, which coincides in the re-im representation with 
% the ordinary gradient wrt u and v
% - the Chain Rule in CR holds and when a holomorphic C-C function is backproped
% the autodiff just wastes cycles on computing its conj-derivative (rather than assuming
% it to be zero)

% Although $\cplx$-valued dense layers and back-propagation have been around since early
% 1990-s, one of the earliest accounts
Development of deep $\cplx$-valued networks has been active. \citet{haensch_complex-valued_2010}
put forward $\cplx$-valued convolutional networks, \citet{guberman_complex_2016} and
\citet{popa_complex-valued_2017} developed modifications of pooling, \citet{arjovsky_unitary_2016}
and \citet{wisdom_full-capacity_2016} proposed $\cplx$-valued RNNs with unitary recurrent
transition matrices, and \citet{danihelka_associative_2016} developed $\cplx$-valued
holographic representations for LSTMs. More recently \citet{trabelsi_deep_2018} proposed
$\cplx$-valued batch-normalization and weight initialization, \citet{wolter_complex_2018}
investigated different $\cplx$-valued gating mechanisms for RNNs, \citet{yang_complex_2019}
proposed $\cplx$-valued self-attention and complex transformer architecture. It merits
noting that \citet{gaudet_deep_2018} generalized $\cplx$VNN further to deep quaternion-valued
networks, and \citet{vecchi_compressing_2020} studied sparsity inducing regularizers for
them.
% complex-valued distributions \citep{pav_moments_2015,taubock_complex-valued_2012},
% and \citep{karseras_caution:_2014}

% This is too abrupt: first intro into what Cplx- valued nets are, then the representation 
% and how it dictates the architecture, and ultimately bioils down to representation
% u(x, y) + j v(x, y) as a R2->R2 map.

% \todo{read on holomorphic nets}
